---
title: "RNA-seq Analysis — Differential Expression Report"
subtitle: "ultimaWTExp"
date: today
format:
  html:
    theme: cosmo
    toc: true
    embed-resources: true
params:
  data_dir: "data"
  outdir: "results"
  run_id: "NA"
execute:
  echo: false
  warning: false
---

```{r setup}
data_dir <- params$data_dir
ls_data  <- if (dir.exists(data_dir)) list.files(data_dir, full.names = TRUE) else character(0)
de_files <- ls_data[grepl("deseq2_results", basename(ls_data))]
```

Audience: **researchers** interpreting the comparisons. Each contrast below lists
the most significant genes and points to its diagnostic plots and enrichment
tables in the results tree.

```{r per-contrast, results='asis'}
if (!length(de_files)) {
  cat("_No differential-expression results were generated for this run._\n")
} else for (f in de_files) {
  cid <- sub("\\.deseq2_results.tsv$", "", basename(f))
  d <- tryCatch(read.delim(f, check.names = FALSE), error = function(e) NULL)
  if (is.null(d)) next
  padj <- if ("padj" %in% names(d)) d$padj else rep(NA, nrow(d))
  lfc  <- if ("log2FoldChange" %in% names(d)) d$log2FoldChange else rep(NA, nrow(d))
  sig  <- !is.na(padj) & padj < 0.05
  cat("\n## ", cid, "\n\n")
  cat(sprintf("- Tested: **%d** genes; significant (padj<0.05): **%d** (up %d / down %d)\n\n",
              nrow(d), sum(sig), sum(sig & lfc > 0, na.rm = TRUE), sum(sig & lfc < 0, na.rm = TRUE)))
  top <- head(d[order(padj), ], 15)
  print(knitr::kable(top, digits = 4, caption = paste("Top genes —", cid)))
  cat("\n\nDiagnostics: `differential_expression/deseq2/", cid,
      "/` (PCA, MA, volcano, dispersion, sample-distance).\n", sep = "")
}
```

## Interpreting the plots

- **PCA** — global structure; condition should drive separation. Batch-driven
  PC1 suggests adding `--blocking_factors batch`.
- **MA / volcano** — effect size vs significance; symmetric volcano = healthy.
- **Dispersion** — DESeq2 fit sanity check; points should hug the fitted curve.
- **Sample-distance heatmap** — replicate concordance and outlier detection.

::: {.callout-note}
Functional interpretation (GO / KEGG / Reactome ORA, optional GSEA) for these gene
lists is under `enrichment/`. Run id: `r params$run_id`.
:::
