---
title: "RNA-seq Analysis — Executive Summary"
subtitle: "ultimaWTExp"
date: today
format:
  html:
    theme: cosmo
    toc: true
    embed-resources: true
params:
  data_dir: "data"
  outdir: "results"
  run_id: "NA"
execute:
  echo: false
  warning: false
---

```{r setup}
data_dir <- params$data_dir
ls_data  <- if (dir.exists(data_dir)) list.files(data_dir, full.names = TRUE) else character(0)
read_if  <- function(pat) {
  f <- ls_data[grepl(pat, basename(ls_data))]
  if (length(f)) tryCatch(read.delim(f[1], check.names = FALSE), error = function(e) NULL) else NULL
}
```

## Purpose

This is the high-level summary for **Principal Investigators, project managers and
collaborators**. It reports what was measured, the headline differential-expression
findings, and where to find the detailed evidence. Technical depth lives in the
companion *Technical* and *QC* reports.

## What was run

ultimaWTExp processed the submitted FASTQ files end to end: read QC and trimming,
alignment, gene/transcript quantification, differential expression, and functional
interpretation. Every step's exact software version is recorded in
`pipeline_info/software_versions.yml`, and all primary deliverables are checksummed
in `SHA256SUMS.txt` for reproducibility.

## Headline results

```{r headline, results='asis'}
de_files <- ls_data[grepl("deseq2_results", basename(ls_data))]
if (length(de_files) == 0) {
  cat("No differential-expression tables were produced (no design / contrasts).\n")
} else {
  for (f in de_files) {
    d <- tryCatch(read.delim(f, check.names = FALSE), error = function(e) NULL)
    if (is.null(d)) next
    cid <- sub("\\.deseq2_results.tsv$", "", basename(f))
    padj <- if ("padj" %in% names(d)) d$padj else rep(NA, nrow(d))
    sig  <- sum(!is.na(padj) & padj < 0.05)
    cat(sprintf("- **%s**: %d genes tested, **%d significant** (padj < 0.05).\n",
                cid, nrow(d), sig))
  }
}
```

## Where to look next

| Question | Deliverable |
|---|---|
| Are the libraries good quality? | `multiqc/multiqc_report.html`, *QC report* |
| Which genes changed? | `differential_expression/`, *Differential report* |
| What biology is implicated? | `enrichment/` (GO / KEGG / Reactome) |
| How were results produced? | `pipeline_info/`, *Technical report* |

::: {.callout-note}
Generated automatically by ultimaWTExp. Run id: `r params$run_id`.
:::
