---
title: "RNA-seq Analysis — QC Report"
subtitle: "ultimaWTExp"
date: today
format:
  html:
    theme: cosmo
    toc: true
    embed-resources: true
params:
  data_dir: "data"
  outdir: "results"
  run_id: "NA"
execute:
  echo: false
  warning: false
---

```{r setup}
data_dir <- params$data_dir
ls_data  <- if (dir.exists(data_dir)) list.files(data_dir, full.names = TRUE) else character(0)
```

Audience: **sequencing facilities and data managers**. This report orients you to
the quality-control evidence; the interactive detail is in the MultiQC report.

## MultiQC

The aggregated, interactive QC report is the primary artefact:

> **`multiqc/multiqc_report.html`**

It consolidates FastQC (raw + trimmed), trimming logs, STAR/HISAT2 alignment
summaries, samtools stats/flagstat/idxstats, Picard duplication and RNA-seq
metrics, Qualimap, RSeQC, preseq and dupRadar — plus the exact software versions.

## Metrics covered

| Stage | Metric | Tool |
|---|---|---|
| Reads | counts, base quality, GC, adapters, duplication, overrepresented seqs | FastQC |
| Trimming | adapter/quality/polyG-X removal | fastp / Trim Galore / Cutadapt |
| Alignment | mapping rate, multimapping | STAR / HISAT2, samtools |
| Duplication | optical/PCR dup rate, dup-vs-expression | Picard, dupRadar |
| Composition | exonic/intronic/intergenic, 5'→3' bias | Picard CollectRnaSeqMetrics, Qualimap |
| Coverage | gene-body coverage uniformity | RSeQC |
| Strandedness | inferred protocol | RSeQC infer_experiment |
| Junctions | junction saturation / annotation | RSeQC |
| Complexity | library saturation curve | preseq |

```{r files, results='asis'}
if (length(ls_data)) {
  cat("\n### Artefacts staged for this report\n\n")
  for (f in basename(ls_data)) cat("- `", f, "`\n", sep = "")
}
```

::: {.callout-tip}
A sample failing several FastQC modules, with low mapping rate or strong 3' bias,
is a candidate for re-prep or exclusion — flag it back to the submitting lab.
:::
