---
title: "RNA-seq Analysis — Technical Report"
subtitle: "ultimaWTExp"
date: today
format:
  html:
    theme: cosmo
    toc: true
    code-fold: true
    embed-resources: true
params:
  data_dir: "data"
  outdir: "results"
  run_id: "NA"
execute:
  echo: false
  warning: false
---

```{r setup}
data_dir <- params$data_dir
ls_data  <- if (dir.exists(data_dir)) list.files(data_dir, full.names = TRUE) else character(0)
```

Audience: **bioinformaticians, data scientists, geneticists**. This report documents
methods, parameters and intermediate artefacts in enough detail to reproduce or
audit the analysis.

## Pipeline

```
FASTQ ─▶ FastQC ─▶ trim (fastp/Trim Galore/Cutadapt) ─▶ FastQC
      ─▶ align (STAR 2-pass / HISAT2) ─▶ samtools + Picard + Qualimap + RSeQC + preseq + dupRadar
      ─▶ quantify (featureCounts / HTSeq / STAR + Salmon→tximport)
      ─▶ DESeq2 (+edgeR) ─▶ clusterProfiler (ORA/GSEA)
      ─▶ STAR-Fusion, rMATS ─▶ MultiQC + reports
```

## Count matrices

```{r matrices, results='asis'}
for (m in c("raw", "tpm", "fpkm", "cpm")) {
  f <- ls_data[grepl(paste0("counts\\.", m, "\\.tsv$"), basename(ls_data))]
  if (length(f)) {
    d <- tryCatch(read.delim(f[1], check.names = FALSE), error = function(e) NULL)
    if (!is.null(d)) cat(sprintf("- **%s**: %d genes × %d samples (`%s`)\n",
                                 toupper(m), nrow(d), ncol(d) - 1, basename(f[1])))
  }
}
```

## Differential expression — methods

DE was computed with **DESeq2** (negative-binomial GLM, Wald test, BH-adjusted).
The design formula adds any `--blocking_factors` (e.g. `batch`, `subject_id`)
ahead of the variable of interest, so multifactor, paired and batch-corrected
models are handled by the same code path. The reference level sets the
fold-change denominator. When `--run_edger` is set, edgeR's quasi-likelihood
F-test provides an orthogonal cross-check.

```{r de-tables, results='asis'}
de_files <- ls_data[grepl("deseq2_results", basename(ls_data))]
if (!length(de_files)) cat("_No DE tables present._\n")
for (f in de_files) {
  cid <- sub("\\.deseq2_results.tsv$", "", basename(f))
  d <- tryCatch(read.delim(f, check.names = FALSE), error = function(e) NULL)
  if (is.null(d)) next
  cat("\n### ", cid, "\n\n")
  top <- head(d[order(d$padj), ], 10)
  print(knitr::kable(top, digits = 4))
}
```

## Reproducibility

- Exact tool versions: `pipeline_info/software_versions.yml`
- Execution trace / timeline / DAG: `pipeline_info/`
- Deliverable checksums: `SHA256SUMS.txt`
- Containers pin the entire toolchain; `-resume` re-uses cached steps.

::: {.callout-note}
Run id: `r params$run_id`.
:::
